---
title: "Chapter 3 - A review of R modeling fundamentals"
author: "Brad Hill"
date: "8/12/2021"
output: html_document
---


```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
data(crickets, package = 'modeldata')
```

## Section 3 - Why tidiness is important?  
This section mostly talks about how different modeling frameworks use different defaults and return counter-intuitive results. It gives us some ways around some of these as well as introducing `broom::tidy()` as a way of cleaning several objects.

```{r}
corr_res <- mtcars %>% 
  select(-mpg) %>% 
  map(cor.test, y = mtcars$mpg)
corr_res[[1]]
```

The above output is not really useable for modeling or visualization purposes. It tells us plenty of information, but not in a very parse-friendly way.

```{r}
broom::tidy(corr_res[[1]])
```

`broom` does a good job of allowing us to use that in a nice easy way.

```{r}
corr_res %>% 
  map_dfr(tidy, .id = 'predictor') %>% 
  ggplot(aes(x = fct_reorder(predictor, estimate))) +
  geom_point(aes(y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  labs(x = '',
       y = 'Correlation with mpg') +
  theme_bw()
```

## Section 4 - Combining Base R and Tidyverse  

We can do a lot simply by combining base R modeling functionality with tidyverse. For instance, modeling separately based on cricket species.

```{r}
(split_by_species <- crickets %>% 
  group_nest(species))

(model_by_species <- split_by_species %>% 
    mutate(model = map(data, ~lm(rate ~ temp, data = .x))))

model_by_species %>% 
  mutate(coef = map(model, tidy)) %>% 
  select(species, coef) %>% 
  unnest(cols = c(coef))
```

